\vspace{-0.25em}\thetitle

### 2009-07-28 paper.tex
...

% At this point, it may be wise to address what we mean here by prediction.
% Brief reflection shows that one cannot hope to accurately predict the exact behavior of flows:
% flow behavior is nondeterministic---flows may exhibit the same initial behavior but subsequently behave quite differently.
% The best we can do then is to predict a \emph{distribution} of possible outcomes given the observed initial behavior.
% When compared with a specific outcome, such a prediction cannot be said to be ``wrong'' or ``right.''
% How then can one measure the quality of a predicted distribution of outcomes?
% This subtle and difficult issue is central to our evaluation methodology.

@@

\Figure{svd} shows scatter plots of most significant two coordinates when this procedure is applied to a random sample of 5000 flows from each of six unrelated traffic traces.
Clear linear structure is visible in all six traces:
flow behaviors are heavily concentrated near low-dimensional subspaces.
Behaviors concentrated along a one-dimensional subspace appear as dark ``spots'' where many points have been plotted;
behaviors near two-dimensional subspaces appear as lines;
behaviors on higher-dimensional subspaces appear as ``smears.''

@@

The properties of the model exactly explain the structures seen in \Figure{svd}.
% This model is an ideal, and observed behaviors have noise introduced by sampling, imperfect adherence to the model, and outliers.
However, this is only one possible hypothesis that fits the data.
Like any hypothesis, it must be tested.
The rest of this paper, aside from providing a practical application, serves as a hypothesis test:
we try to recover the matrices $W$ and $P$ from our noisy and imperfect observations of $Q$ and use the recovered model to predict real flow behaviors.
If the recovered model can make accurate predictions, this provides evidence that our model and hypothesis are valid.

@@

Our experimental methodology has three parts:
training, prediction and evaluation.
For training, we take a sample of network trace traffic and try to recover the structure hypothesized in \Section{hypothesis} from it.
For prediction, we use the trained model to predict behaviors of a separate set of flows from the same trace, using only knowledge of a few initial packets of each.
Finally, we must evaluate the quality of these predictions.
As noted in the \Section{introduction}, due to nondeterminism we cannot possibly predict exact behaviors of flows and must instead evaluate how good a predicted distribution of possible behaviors is as compared to actual outcomes.
% One relatively intuitive measure of prediction quality we use is threshold testing.

% Our methodology for recovering the hypothesized linear structure of network traffic from trace samples uses a variety of powerful computational tools:
% \begin{enumerate}
%   \item Compute the linear representation matrix of the trace.
%   \item Perform linear subspace segmentation, extracting subspaces near which significant amounts of data points lie.
%   % For this step, we use Ma~\emph{et~al.}'s 2007 algorithm based on lossy compression of data~\cite{Ma07}.
%   \item Use \caps{SVD} to determine an approximation rank for each segmentation and an optimal approximation basis.
%   \item Compute the vertices of the intersection of each approximation subspace with the standard simplex.
%   For each subspace, all nonnegative points can be written as convex linear combinations of these vertices.
%   % For this we use Fukuda's excellent \caps{CDD}lib software for high-dimensional polyhedral computations, implementing a variety of convex hull and vertex enumeration algorithms~\cite{Avis92}.
%   \item Find optimal weights to approximate each data point from the vertices of its subspace.
% \end{enumerate}

@@

We embed the following features of flow behavior into vector spaces via feature-frequency linear representation:
packet size distribution,
inter-packet interval distribution,
packet count,
\caps{IP} protocol type,
\caps{TCP/UDP} port numbers.
All features are represented using frequency histograms.
For packet size distribution, this is trivial since packet sizes are discrete, taking on only values from 1 to the \caps{MTU}, which was 1500 in all of our traces.
Inter-packet intervals must be quantized, using the scheme described by Karpinski~\emph{et~al.}~\cite{Karpinski08} but using only 500 dimensions instead of 1500.
This scheme maps bins of interval durations to indices, with the bins starting at the scale of milliseconds, and ramping smoothly up to seconds at the high end of the scale.
Intervals of more than ten minutes are considered to start a new flow.

\caps{IP} protocol type is naturally discrete, with values from 0 to 255, represented by a 256-dimensions histogram vector.
The ``histogram'' for a single flow is always just an indicator vector with a single unit entry.
However, the linear embedding of this feature into a frequency space allows us to infer the distribution of possible values using linear analysis.
Similarly, port numbers are naturally discrete.
However, the range of port numbers (0-65535) is too large for practical representation using a full histogram.
Instead, we represent port numbers by their remainder modulo 1000.
Port numbers are sufficiently spread out that this still provides ample distinction between traffic types.
We also represent the packet count of each flow.
Since the potential packet count is unlimited, we must take some approach to limiting the dimension of this frequency representation.
We use the simplest approach and simply cap the packet count at 3000 packets:
all flows with 3000 or more packets fall into the same histogram bin.
In practice, flows longer than 3000 packets are ``long.''

Each flow is thus represented as a direct sum of frequency vectors:
\begin{math}
  \text{size} \directsum
  \text{ival} \directsum
  \text{type} \directsum
  \text{ports} \directsum
  \text{pkts} \in \R^{6256}.
\end{math}
A collection of flows is represented as a ``stacked'' matrix of row vectors:
\begin{align}
  X = [ ~
  \text{Size} ~
  \text{Ival} ~
  \text{Type} ~
  \text{Ports} ~
  \text{Pkts} ~
  ] \in \R^{m \times 6256}.
\end{align}

>>>>

@@

\Figure{svd} shows scatter plots of most significant two coordinates when this procedure is applied to a random sample of 5000 flows from each of six independent traffic traces.
Clear linear structure is visible in all six traces:
flow behaviors are heavily concentrated near low-dimensional subspaces.
Behaviors concentrated along a one-dimensional subspace appear as dark ``spots'' where many points have been plotted;
behaviors near two-dimensional subspaces appear as lines;
behaviors on higher-dimensional subspaces appear as ``smears.''

@@

This model explains the structures found in \Figure{svd}.
Points lying along the same structure are in the same class, and each structure is ``generated'' by a small set of vertices:
the points belonging to a structure are near the convex hull generated by those vertices.
However, this is only one possible hypothesis that fits the data.
Like any hypothesis, it must be tested.
The rest of this paper, aside from providing a practical application, serves as a hypothesis test:
we try to recover the matrices $W$ and $P$ from our noisy and imperfect observations of $Q$ and use the recovered model to predict real flow behaviors.
If the recovered model can make accurate predictions, this provides evidence that our model and hypothesis are valid.

@@

Our experimental methodology has three parts:
training, prediction and evaluation.
For training, we take a sample of network trace traffic and try to recover the model hypothesized in \Section{hypothesis} from it.
Next, we use the trained model to predict behaviors of a separate set of flows from the same trace, using only knowledge of a few initial packets of each flow.
Finally, we must evaluate the quality of these predictions by comparing them to the actual behavior of those flows.

It seems wise at this point to address the nature of prediction for highly nondeterministic phenomena like network traffic.
Brief reflection shows that one cannot hope to accurately predict the exact behavior of flows:
flows may exhibit the same initial behavior but subsequently behave quite differently.
The best we can do then is to predict a \emph{distribution} of possible outcomes given the observed initial behavior.
When compared with a specific outcome, such a prediction cannot be said to be ``wrong'' or ``right.''
How then can one measure the quality of a predicted distribution of outcomes?
This subtle and difficult issue is central to our evaluation methodology.

@@

We embed the following features of flow behavior into vector spaces via feature-frequency representation:
packet size distribution,
inter-packet interval distribution,
\caps{IP} protocol type,
\caps{TCP/UDP} port numbers,
and packet count.
Packet sizes are already naturally discrete, taking on only values from 1 to the \caps{MTU}, which was 1500 in all of our traces.
Inter-packet intervals must be quantized, using the scheme described by Karpinski~\emph{et~al.}~\cite{Karpinski08} but using only 500 dimensions instead of 1500.
This scheme maps bins of interval durations to indices, with the bins starting at the scale of milliseconds, and ramping smoothly up to seconds at the high end of the scale.
Intervals of more than ten minutes are considered to start a new flow.
\caps{IP} protocol type also naturally discrete, with values from 0 to 255, represented by a 256-dimensions histogram vector.
The feature-frequency vector for a single flow in this case is always just a vector with a single unit entry and all other entries zero.
Port numbers are naturally discrete as well, but the range of port numbers (0-65535) is too large for practical representation using a full histogram.
Instead, we represent port numbers by their remainder modulo 1000.
Port numbers are sufficiently spread out that this still provides ample distinction between traffic types.
Moreover, since the direction of a flow is generally uninteresting, we represent both source and destination in a single vector containing two values.
We also represent the packet count of each flow.
Since the potential packet count is unlimited, some approach is needed for limiting the dimension of this frequency representation.
We use the simplest approach and simply cap the packet count at 3000 packets:
all flows with 3000 or more packets fall into the same histogram bin and are simply considered ``very large.''
